Stack Safety as a Security Property

### 2020-07-09 main.tex
pass on contributions

\bcp{We don't actually say very clearly anywhere that we are interested in
  low-level enforcement mechanisms for machine-code programs (because this
  is where stack attacks occur, and because at higher levels of abstraction
  the stack is a built-in concept, not something that can be attacked even
  in principle).  This could cause confusion.}

@@

But what {\em is} this stack safety of which people speak? Precise formal
definitions are mostly lacking: rather, the notion
is commonly defined in a negative way, by enumerating the bad things that a
given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables\bcp{Does use of unitialized variables lead to
stack safety attacks?}.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a
positive definition, by
\citet{Skorstengaard+19}, who define stack safety as
the conjunction of two properties: {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this represents an important step in the right direction, we believe
there is still room to craft a more abstract, more generic characterization
of stack safety that is independent of capabilities or any other specific
enforcement
technique.  In other words, we seek formal criteria that can be applied
to \emph{any} mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
%
Our key contribution is a novel, generic presentation of
local state encapsulation and well-bracketed control flow using the
technical framework of language-based security.

@@

The intuition behind \citeauthor{Skorstengaard+19}'s {\em well-bracketed control
  flow} is
the need to protect the call graph itself: a callee should pass control
back to the caller only by returning in the standard way (transferring
control to the instruction following the call); it should
not be
allowed to jump back into the caller's code at an arbitrary point, either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety
between functions: each instruction is uniquely owned by some
function, and all jumps between functions are either calls that enter
at a marked entry point or returns that re-enter just after their
associated call in the standard nested call structure.  \apt{clarify
  what we are contributing here.}

To phrase these two aspects of stack safety as security properties requires
an explicit attacker model: a clear understanding of who is being protected
and from what. Most existing stack protection mechanisms envisage scenarios
where a callee is influenced by external inputs to corrupt the stack in some
way. To avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we state the
protections that a function can rely on when it makes a call, we assume that
the callee can execute arbitrary instructions until it returns.

@@

After formally defining stack safety, we show that our definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
conservative Depth Isolation policy, which not only enforces the
our formal definition of stack safety, but also a much stronger
property that requires integrity and confidentiality at every step of
an execution trace. This property is provably (in Coq)\leo{How to
  mention this?} stronger and stemmed out of the need to more
efficiently test stack safety. However, at the same time the
conservative enforcement mechanism is expensive to implement. In the
same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different
mechanism that is more efficient by virtue of being lazy in its
enforcement. At a high level, instead of guaranteeing integrity and
confidentiality when a callee returns, they check it when the caller
actually makes use of (potentially compromised) data. Thus, they
enforce a (also provably in Coq) weaker property that still protects
from unsafe callees, but is much less intuitive.

In summary, we offer the following contributions:\bcp{I would still like to
  see these points phrased in ``contribution normal form'': say exactly what
  we did for each one (it's a bit vague---''considers,'' etc.), and
  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a generic setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our strict stack safety property, and also a (provably in Coq) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our strict
  local state encapsulation property,
  but a modified version of it using unique per-activation tags enforces
  an (also provably) weaker lazy variant of the property, which still prohibits
  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer to throughout the paper
(\cref{sec:running-example}), a description of our threat model
(\cref{sec:threat}) and some technical preliminaries
(\cref{sec:prelim}).

@@

\subsection{A Conservative Policy}
%
The first micropolicy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. Beyond
those capabilities, the compiler generates designated sequences of
instructions with special tags that grant them permission to manage
the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\subsection{A Stronger Lockstep Stack Safety Property}

>>>>

\bcp{We don't actually say very clearly anywhere that we are interested in
  low-level enforcement mechanisms for machine-code programs (because this
  is where stack attacks occur, and because at higher levels of abstraction
  the stack is a built-in concept, not something that can be attacked even
  in principle).  This could cause confusion.\apt{not so worried}}

@@

But what {\em is} this stack safety of which people speak? Precise formal
definitions are mostly lacking: rather, the notion
is commonly defined in a negative way, by enumerating the bad things that a
given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables\bcp{Does use of unitialized variables lead to
stack safety attacks?\apt{sure, via buffer overflow}}.
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a
positive definition, by
\citet{Skorstengaard+19}, who define stack safety as
the conjunction of two properties: {\em local state encapsulation} and {\em
  well-bracketed control flow}.
They formalize these properties by defining an abstract capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this represents an important step in the right direction, we believe
there is still room to craft a more abstract, more generic characterization
of stack safety that is independent of capabilities or any other specific
enforcement
technique.  In other words, we seek formal criteria that can be applied
to \emph{any} mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
%
Our key contribution is a novel, generic presentation of
local state encapsulation and well-bracketed control flow using the
technical framework of language-based security.

@@

The intuition behind \citeauthor{Skorstengaard+19}'s {\em well-bracketed control
  flow} is
the need to protect the call graph itself: a callee should pass control
back to the caller only by returning in the standard way (transferring
control to the instruction following the call); it should
not be
allowed to jump back into the caller's code at an arbitrary point, either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety
between functions: each instruction is uniquely owned by some
function, and all jumps between functions are either calls that enter
at a marked entry point or returns that re-enter just after their
associated call in the standard nested call structure.  \apt{clarify
  what we are contributing here. We are not trying to phrase this as a security property.}

To phrase these two \apt{which two? Not WBCF!} aspects of stack safety as security properties requires
an explicit attacker model: a clear understanding of who is being protected
and from what. Most existing stack protection mechanisms envisage scenarios
where a callee is influenced by external inputs to corrupt the stack in some
way. To avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we state the
protections that a function can rely on when it makes a call, we assume that
the callee can execute arbitrary instructions until it returns.

@@

After formally defining stack safety, we show that our definition is
enforceable using an existing tag-based enforcement
mechanism~\citep{DBLP:conf/sp/RoesslerD18}. We first consider their
conservative Depth Isolation policy, which enforces the
our basic stack safety property. It also enforces an alternative
lockstep property that requires integrity and confidentiality at every step of
an execution trace, which fails faster and thus is more efficient to test;
we prove (in Coq) that the lockstep property is stronger 
than our basic property,
However, at the same time the
conservative enforcement mechanism is expensive to implement. In the
same work, \citeauthor{DBLP:conf/sp/RoesslerD18} propose a different
mechanism that is more efficient by virtue of being lazy in its
enforcement. At a high level, instead of guaranteeing integrity and
confidentiality when a callee returns, this policy checks it when the caller
actually makes use of (potentially compromised) data. The lazy policy
does not enforce our basic propety, but an enhanced version of it does
enforce a proveably weaker property that still protects
from unsafe callees, but is less intuitive.

In summary, we offer the following contributions:\bcp{I would still like to
  see these points phrased in ``contribution normal form'': say exactly what
  we did for each one (it's a bit vague---''considers,'' etc.), and
  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  A novel formal stack safety property capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property is enforced by 
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:conservative}).
\item
  A lockstep stack safety property that is also enforced by the Depth Isolation policy,
  and is more efficient to test, which we prove (in Coq) is stronger than the basic property(~\cref{sec:lockstep}).
\item
  A novel lazy stack safety property that is enforced by the more efficient
  Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}, which we prove (in Coq) is weaker than the basic property~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.
%
We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).

\newcommand*{\rsp}{\textsc{sp}}

@@

\subsection{A Conservative Policy}
\label{sec:conservative}
%
The first micropolicy we present to enforce the definition of stack
safety developed in \cref{sec:lse-and-wbcf} is a small variation
on the Depth Isolation policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the stack memory
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. Beyond
those capabilities, the compiler generates designated sequences of
instructions with special tags that grant them permission to manage
the stack, say during calls and returns.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\subsection{A Stronger Lockstep Stack Safety Property}
\label{sec:lockstep}
